### 2022-01-11 paper/paper.tex
move FMA discussion to the appendix

As demonstrated in Section~\ref{sec:performance-comparison},
Julia is not generically slower than traditional high-performance programming
languages like C, C++, and Fortran. A minor exception in the context of scientific
computing relying on floating point operations is that Julia does not enable
fused multiply-add (FMA) contraction by default, while GCC and Intel compilers
do\footnote{An interesting coincidence is that not only Julia but most LLVM-based
frontends tend not to enable FMA contraction by default.}.
Thus, one must opt-in to these optimizations explicitly.
Relying on Julia's code manipulation techniques, this can be achieved with the macro
\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}},
which makes it convenient to use Julia's function \lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}.

@@

>>>>

As demonstrated in Section~\ref{sec:performance-comparison},
Julia is not generically slower than traditional high-performance programming
languages like C, C++, and Fortran. A minor exception in the context of scientific
computing relying on floating point operations is that Julia does not perform
automatic fused multiply-add (FMA) contraction, which can be remedied by
explicitly using the
\lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
function or the more convenient
\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
macro (for more details see the appendix).

@@

\appendix

Julia does not perform automatic fused multiply-add (FMA) contraction, that is
replacing a multiplication followed by an addition with a single FMA instruction,
as it is inconsistent with its strict floating point semantics.
In other compilers, this optimization is controlled by a global compiler option
(\lstinline{-ffp-contract} in GCC and Clang) and often enabled at reasonable
optimization levels such as \lstinline{-O2} for GCC and Intel compilers\footnote{%
An interesting coincidence is that not only Julia but most LLVM-based frontends
tend not to enable FMA contraction by default.}.
In Julia, FMA contraction is controlled more locally: the
\lstinline{muladd}%\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
intrinsic function allows the compiler to evaluate a multiply-add using the most
efficient method. Relying on Julia's code manipulation techniques, the
\lstinline{@muladd}%\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
macro makes it convenient to use the \lstinline{muladd} function by
syntactically rewriting expressions to insert \lstinline{muladd} in appropriate
locations.
